# Why PLS doesn’t always work

## Learning goals

-   Understand why PLS regression sometimes doesn’t perform as expected
-   Be able to make temperature response plots for agroclimatic metrics
-   Be able to relate temperatures at a study site to temperature responses of models to anticipate whether PLS analysis will be effective

## Disappointing PLS performance

PLS regression makes it difficult to distinguish temperature response phases, especially during the chilling period. Replacing temperatures with agroclimatic metrics did not fully resolve this issue. Expected patterns often appeared only at the beginning and end of the presumed endodormancy phase, with a large gap in between. To understand this, it is important to consider the limitations of the applied analysis method.

## What PLS regression can find

PLS regression identifies relationships between a target variable and variations in input variables, requiring a monotonic response, meaning the correlation must remain consistently positive or negative. Additionally, successful PLS analysis depends on sufficient variation in the independent variables, as meaningful responses cannot be detected without it.

In this case, a constant level of chill effectiveness over a certain period may prevent PLS regression from producing useful results. This could explain its limited performance, particularly in cold regions such as Klein-Altendorf or Beijing, where winter chill accumulation shows little variation.

To investigate this, the model’s response to temperature will be visualized. The analysis is based on the Dynamic Model and examines its reaction to daily temperature patterns. Instead of single-day calculations, the average daily chill accumulation rate over extended periods will be considered, as the model often requires multiple days to accumulate a Chill Portion.

```{r}
library(chillR)

mon <- 1 # Month
ndays <- 31 # Number of days per month
tmin <- 1
tmax <- 8
latitude <- 50


weather <- make_all_day_table(
            data.frame(Year = c(2001, 2001),
                       Month = c(mon, mon),
                       Day = c(1, ndays),
                       Tmin = c(0, 0),
                       Tmax = c(0, 0))) %>%
  mutate(Tmin = tmin,
         Tmax = tmax)

hourly_temps <- stack_hourly_temps(weather,
                                   latitude = latitude)

CPs <- Dynamic_Model(
     hourly_temps$hourtemps$Temp)

daily_CPs <- CPs[length(CPs)] / nrow(weather)

daily_CPs

```

To introduce more flexibility into the code, an alternative way to execute a function is used. Previously, the `Dynamic_Model` function was called directly, but to allow different models to be used dynamically, the `do.call` function is applied. This approach enables passing the model as a parameter, making it possible to switch between models like Dynamic_Model and GDH by assigning them to the `temp_model` variable. The `list(...)` inside `do.call` contains the function arguments, ensuring adaptability.

So far, calculations were performed for a single Tmin/Tmax combination and only for one month. To extend the analysis, data will be generated for all months and a broader range of Tmin/Tmax values. This requires more complex code to handle generalization challenges, such as determining the number of days in each month.

```{r, eval=FALSE}
library(chillR)

latitude <- 50.6

month_range <- c(10, 11, 12, 1, 2, 3)

Tmins <- c(-20:20)
Tmaxs <- c(-15:30)

mins <- NA
maxs <- NA
CP <- NA
month <- NA
temp_model <- Dynamic_Model

for(mon in month_range)
    {days_month <- as.numeric(
      difftime( ISOdate(2002,
                        mon + 1,
                        1),
                ISOdate(2002,
                        mon,
                        1)))
    
    if(mon == 12) days_month <- 31
    
    weather <- make_all_day_table(
                data.frame(Year = c(2001, 2001),
                           Month = c(mon, mon),
                           Day = c(1, days_month),
                           Tmin = c(0, 0),
                           Tmax = c(0, 0)))
    
    for(tmin in Tmins)
      for(tmax in Tmaxs)
        if(tmax >= tmin)
          {
          hourtemps <- weather %>%
            mutate(Tmin = tmin,
                   Tmax = tmax) %>%
            stack_hourly_temps(latitude = latitude) %>% 
            pluck("hourtemps", "Temp")

          CP <- c(CP,
                  tail(do.call(temp_model,
                               list(hourtemps)), 1) /
                    days_month)
          mins <- c(mins, tmin)
          maxs <- c(maxs, tmax)
          month <- c(month, mon)
        }
}

results <- data.frame(Month = month,
                      Tmin = mins,
                      Tmax = maxs,
                      CP)

results <- results[!is.na(results$Month), ]


write.csv(results,
          "data/model_sensitivity_development.csv",
          row.names = FALSE)
```

```{r, eval=FALSE}
head(results)
```

```{r, echo=FALSE }
results <- read.csv("data/model_sensitivity_development.csv")
month_range <- c(10, 11, 12, 1, 2, 3)
library(kableExtra)
kable(head(results), align = c("c")) %>%
  kable_styling("striped", position = "left", font_size=14)
```

```{r}
latitude <- 50.6

month_range <- c(10, 11, 12, 1, 2, 3)

Tmins <- c(-20:20)
Tmaxs <- c(-15:30)
```

With the generated data, the temperature response curve can now be plotted, visualizing how the model reacts to different Tmin/Tmax combinations over time.

```{r, warning=FALSE, message=FALSE}

library(ggplot2)
library(colorRamps)

results$Month_names <- factor(results$Month,
                              levels = month_range,
                              labels = month.name[month_range])  

DM_sensitivity <- ggplot(results,
                         aes(x = Tmin,
                             y = Tmax,
                             fill = CP)) +
  geom_tile() +
  scale_fill_gradientn(colours = alpha(matlab.like(15),
                                       alpha = .5),
                       name = "Chill/day (CP)") +
  ylim(min(results$Tmax),
       max(results$Tmax)) +
  ylim(min(results$Tmin),
       max(results$Tmin))

DM_sensitivity
```

```{r}
DM_sensitivity <- DM_sensitivity +
  facet_wrap(vars(Month_names)) +
  ylim(min(results$Tmax),
       max(results$Tmax)) +
  ylim(min(results$Tmin),
       max(results$Tmin))

DM_sensitivity
```

The vibrant colors in the upper plot result from the `alpha` function in the `ggplot` call, which makes colors transparent. Since data points for six months are layered on top of each other, the overlapping transparency creates more intense colors than those in the legend.

This plot is specific to the given latitude, as daily temperature curves are influenced by sunrise and sunset times. The visualization reveals clear variations in chill effectiveness across the temperature spectrum, with no accumulation at extreme temperatures and a peak in the middle. A steep gradient in chill accumulation rates is also visible near the edges of the effective temperature range.

To assess the likelihood of PLS regression performing well under these climatic conditions, temperatures can be plotted over this diagram. Since the plot corresponds to latitude 50.6°N, temperature data for Klein-Altendorf can be overlaid for further analysis.

```{r, warning=FALSE}
temperatures <- read_tab("data/TMaxTMin1958-2019_patched.csv") %>%
  filter(Month %in% month_range) %>%
  mutate(Month_names =
           factor(Month,
                  levels = c(10, 11, 12, 1, 2, 3),
                  labels = c("October", "November", "December",
                             "January", "February", "March")))

temperatures[which(temperatures$Tmax < temperatures$Tmin),
             c("Tmax", "Tmin")] <- NA

DM_sensitivity +
  geom_point(data = temperatures,
             aes(x = Tmin,
                 y = Tmax,
                 fill = NULL,
                 color = "Temperature"),
             size = 0.2) +
  facet_wrap(vars(Month_names)) +
  scale_color_manual(values = "black",
                     labels = "Daily temperature \nextremes (°C)",
                     name = "Observed at site" ) +
  guides(fill = guide_colorbar(order = 1),
         color = guide_legend(order = 2)) +
  ylab("Tmax (°C)") +
  xlab("Tmin (°C)") + 
  theme_bw(base_size = 15) 
```

Before analyzing the results in detail, the process will be automated to generate data for additional locations where PLS results have been examined. This requires two functions: one for computing model sensitivity data, as this step is time-intensive, and another for plotting the results. This approach ensures consistency and efficiency in comparing temperature response patterns across different locations.

```{r}
Chill_model_sensitivity <-
  function(latitude,
           temp_models = list(Dynamic_Model = Dynamic_Model,
                              GDH = GDH),
           month_range = c(10, 11, 12, 1, 2, 3),
           Tmins = c(-10:20),
           Tmaxs = c(-5:30))
  {
  mins <- NA
  maxs <- NA
  metrics <- as.list(rep(NA,
                         length(temp_models)))
  names(metrics) <- names(temp_models)
  month <- NA
 
  for(mon in month_range)
    {
    days_month <-
      as.numeric(difftime(ISOdate(2002,
                                  mon + 1,
                                  1),
                          ISOdate(2002,
                                  mon,
                                  1) ))
    if(mon == 12) days_month <- 31
    weather <- 
      make_all_day_table(data.frame(Year = c(2001, 2001),
                                    Month = c(mon, mon),
                                    Day = c(1, days_month),
                                    Tmin = c(0, 0),
                                    Tmax = c(0, 0)))

    
    for(tmin in Tmins)
      for(tmax in Tmaxs)
        if(tmax >= tmin)
          {
          hourtemps <- weather %>%
            mutate(Tmin = tmin,
                   Tmax = tmax) %>%
            stack_hourly_temps(
              latitude = latitude) %>%
            pluck("hourtemps",
                  "Temp")
          
          for(tm in 1:length(temp_models))
            metrics[[tm]] <- 
              c(metrics[[tm]],
                tail(do.call(temp_models[[tm]],
                        list(hourtemps)),1)/
                              days_month)
          
          mins <- c(mins, tmin)
          maxs <- c(maxs, tmax)
          month <- c(month, mon)
        }
    }
  results <- cbind(data.frame(Month = month,
                              Tmin = mins,
                              Tmax = maxs),
                   as.data.frame(metrics))
  
  results <- results[!is.na(results$Month),]
}


Chill_sensitivity_temps <-
  function(chill_model_sensitivity_table,
           temperatures,
           temp_model,
           month_range = c(10, 11, 12, 1, 2, 3),
           Tmins = c(-10:20),
           Tmaxs = c(-5:30),
           legend_label = "Chill/day (CP)")
{
  library(ggplot2)
  library(colorRamps)

  cmst <- chill_model_sensitivity_table
  cmst <- cmst[which(cmst$Month %in% month_range),]
  cmst$Month_names <- factor(cmst$Month,
                             levels = month_range,
                             labels = month.name[month_range])  
  
  DM_sensitivity<-
    ggplot(cmst,
           aes_string(x = "Tmin",
                      y = "Tmax",
                      fill = temp_model)) +
    geom_tile() +
    scale_fill_gradientn(colours = alpha(matlab.like(15),
                                         alpha = .5),
                         name = legend_label) +
    xlim(Tmins[1],
         Tmins[length(Tmins)]) +
    ylim(Tmaxs[1],
         Tmaxs[length(Tmaxs)])
  
  temperatures<-
    temperatures[which(temperatures$Month %in% month_range),]
  
  temperatures[which(temperatures$Tmax < temperatures$Tmin),
               c("Tmax", 
                 "Tmin")] <- NA
  
  temperatures$Month_names <-
    factor(temperatures$Month,
           levels = month_range,
           labels = month.name[month_range])  
  
  DM_sensitivity +
    geom_point(data = temperatures,
               aes(x = Tmin,
                   y = Tmax,
                   fill = NULL,
                   color = "Temperature"),
               size = 0.2) +
    facet_wrap(vars(Month_names)) +
    scale_color_manual(values = "black",
                       labels = "Daily temperature \nextremes (°C)",
                       name = "Observed at site" ) +
    guides(fill = guide_colorbar(order = 1),
           color = guide_legend(order = 2)) +
    ylab("Tmax (°C)") +
    xlab("Tmin (°C)") + 
    theme_bw(base_size = 15)

}
  
```

To compare different locations, the model sensitivity data will first be generated for each site. These datasets will be saved to a file to avoid redundant computations, ensuring efficiency by allowing reuse of the data without reprocessing. This will facilitate a direct comparison of temperature response patterns across locations.

```{r, eval=FALSE}

Model_sensitivities_CKA <-
  Chill_model_sensitivity(latitude = 50,
                          temp_models = list(Dynamic_Model = Dynamic_Model,
                                             GDH = GDH),
                          month_range = c(10:12, 1:5))

write.csv(Model_sensitivities_CKA,
          "data/Model_sensitivities_CKA.csv",
          row.names = FALSE)

Model_sensitivities_Davis <-
  Chill_model_sensitivity(latitude = 38.5,
                          temp_models = list(Dynamic_Model = Dynamic_Model,
                                             GDH = GDH),
                          month_range=c(10:12, 1:5))

write.csv(Model_sensitivities_Davis,
          "data/Model_sensitivities_Davis.csv",
          row.names = FALSE)

Model_sensitivities_Beijing <-
  Chill_model_sensitivity(latitude = 39.9,
                          temp_models = list(Dynamic_Model = Dynamic_Model, 
                                             GDH = GDH),
                          month_range = c(10:12, 1:5))

write.csv(Model_sensitivities_Beijing,
          "data/Model_sensitivities_Beijing.csv",
          row.names = FALSE)

Model_sensitivities_Sfax <-
  Chill_model_sensitivity(latitude = 35,
                          temp_models = list(Dynamic_Model = Dynamic_Model,
                                             GDH = GDH),
                          month_range = c(10:12, 1:5))

write.csv(Model_sensitivities_Sfax,
          "data/Model_sensitivities_Sfax.csv",
          row.names = FALSE)

```

```{r, echo=FALSE}
Model_sensitivities_CKA <- read.csv("data/Model_sensitivities_CKA.csv")
Model_sensitivities_Davis <- read.csv("data/Model_sensitivities_Davis.csv")
Model_sensitivities_Beijing <- read.csv("data/Model_sensitivities_Beijing.csv")
Model_sensitivities_Sfax <- read.csv("data/Model_sensitivities_Sfax.csv")
```

Now, temperature patterns for specific months can be compared to the effective ranges for chill and heat accumulation at the four locations. By analyzing long-term weather data, it becomes possible to assess how well each site's climate aligns with the conditions necessary for dormancy processes.

```{r, echo=FALSE, warning=FALSE}
Beijing_weather <- read_tab("data/Beijing_weather.csv")
CKA_temperatures <- read_tab("data/TMaxTMin1958-2019_patched.csv")
Davis_weather <- read_tab("data/Davis_weather.csv")
Sfax_weather <- read_tab("data/Sfax_weather.csv")
```

To proceed, the long-term weather data for the four locations should be downloaded and saved in a `data` subfolder within the working directory.

```{r, echo=FALSE}
library(downloadthis)
Beijing_weather %>% download_this(
    output_name = "Beijing_weather",
    output_extension = ".csv",
    button_label = "Download weather data for Beijing",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )

CKA_temperatures %>% download_this(
    output_name = "TMaxTMin1958-2019_patched",
    output_extension = ".csv",
    button_label = "Download weather data for Klein-Altendorf",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )

Davis_weather %>% download_this(
    output_name = "Davis_weather",
    output_extension = ".csv",
    button_label = "Download weather data for Davis",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )

Sfax_weather %>% download_this(
    output_name = "Sfax_weather",
    output_extension = ".csv",
    button_label = "Download weather data for Sfax",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

The analysis will begin with Beijing, the coldest location, and then progress through Klein-Altendorf, Davis, and Sfax, representing increasingly warmer growing regions. This step-by-step approach allows for a clear comparison of chill accumulation across different climates.

Since `Chill_sensitivity_temps` outputs a `ggplot` object, it can be modified using the `+`notation in ggplot2. Titles will be added to each plot using the `ggtitle` function to improve clarity. Below are the generated plots for chill accumulation, quantified using the Dynamic Model, illustrating how effective temperature ranges vary across these locations.

```{r, warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_Beijing,
                        Beijing_weather,
                        temp_model = "Dynamic_Model",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        legend_label = "Chill per day \n(Chill Portions)") +
  ggtitle("Chill model sensitivity at Beijing, China")

```

```{r warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_CKA,
                        CKA_temperatures,
                        temp_model = "Dynamic_Model",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        legend_label = "Chill per day \n(Chill Portions)") +
  ggtitle("Chill model sensitivity at Klein-Altendorf, Germany")
```

```{r warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_Davis,
                        Davis_weather,
                        temp_model = "Dynamic_Model",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        legend_label = "Chill per day \n(Chill Portions)") +
  ggtitle("Chill model sensitivity at Davis, California")
```

```{r warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_Sfax,
                        Sfax_weather,
                        temp_model = "Dynamic_Model",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        legend_label = "Chill per day \n(Chill Portions)") +
  ggtitle("Chill model sensitivity near Sfax, Tunisia")
```

Below are the plots for heat accumulation:

```{r, warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_Beijing,
                        Beijing_weather,
                        temp_model = "GDH",
                        month_range = c(12, 1:5),
                        legend_label = "Heat per day \n(GDH)") +
  ggtitle("Heat model sensitivity at Beijing, China")
```

```{r warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_CKA,
                        CKA_temperatures,
                        temp_model = "GDH",
                        month_range = c(12, 1:5),
                        legend_label = "Heat per day \n(GDH)") +
  ggtitle("Heat model sensitivity at Klein-Altendorf, Germany")
```

```{r warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_Davis,
                        Davis_weather,
                        temp_model = "GDH",
                        month_range = c(12, 1:5),
                        legend_label = "Heat per day \n(GDH)") +
  ggtitle("Heat model sensitivity at Davis, California")
```

```{r warning=FALSE}
Chill_sensitivity_temps(Model_sensitivities_Sfax,
                        Sfax_weather,
                        temp_model = "GDH",
                        month_range = c(12, 1:5),
                        legend_label = "Heat per day \n(GDH)") +
  ggtitle("Heat model sensitivity near Sfax, Tunisia")
```

Now, the response patterns will be analyzed in relation to the sensitivity of the chill model. This will help identify how different locations align with the effective temperature ranges for chill accumulation and how variations in climate influence dormancy processes.

## Chill model sensitivity vs. observed temperature

Although chill model sensitivity is influenced by hourly temperatures, which depend on sunrise and sunset times, the overall sensitivity patterns appear nearly identical across all locations. While minor differences exist, they are not significant enough to impact the broader interpretation.

What does vary, however, is how observed hourly temperatures align with the sensitive periods for chill accumulation. This alignment shifts not only between locations but also across different months, highlighting the importance of local climatic conditions in determining the effectiveness of chill accumulation.

###  Beijing, China

In Beijing, the relationship between observed temperatures and effective chill accumulation varies significantly across months:

-   **October:** Temperatures are evenly distributed between effective chill accumulation conditions and values that are too warm. This variation may generate **chill-related signals**, potentially influencing **bloom dates** in a detectable way.

-   **November:** Almost all temperatures fall within the highly effective chill accumulation range. Despite considerable temperature fluctuations, chill accumulation rates remain similar for most days. Without meaningful variation, PLS regression is unlikely to produce useful results.

-   **December–February:** Temperatures drop further, leading to many hours below the effective chill accumulation range. If the Dynamic Model correctly represents chill accumulation at low temperatures (which remains uncertain), a response might still be observed.

-   **March:** Similar to **November**, temperatures remain almost always optimal for chill accumulation, resulting in limited variation.

**Summary:**

Beijing's conditions should allow for clear identification of chill effects in most months, except for November, where the lack of variation may reduce the effectiveness of PLS regression.

### Klein-Altendorf, Germany

Compared to Beijing, Klein-Altendorf appears to be less suitable for using PLS regression to delineate the chilling period. While both locations experience a mix of optimal and suboptimal temperatures, the proportion of days with suboptimal chill conditions is much lower in Klein-Altendorf.

-   **October, December, January, February:** Most days have near-optimal chill conditions, with only a few too warm days (mainly in October) or too cold days for effective chill accumulation.

-   **November and March:** Almost all days fall within the optimal chill accumulation range, leading to minimal variation.

Summary:

Due to the lack of strong temperature fluctuations, Klein-Altendorf may not be well-suited for PLS regression analysis, as the method relies on meaningful variation to detect temperature response phases.

### Davis, California

Davis provides more favorable conditions for PLS regression compared to the colder locations, particularly due to the distribution of daily temperatures across chill model sensitivity levels during certain months.

-   **November and March:** Daily temperatures are evenly distributed between effective and ineffective chill accumulation conditions, creating meaningful variation that could help detect temperature response phases.

-   **February:** Most days have optimal chill accumulation conditions, with relatively few days experiencing low or zero chill accumulation.

-   **January and December:** Conditions are mostly optimal for chill accumulation, with little variation, which may limit PLS regression’s ability to detect meaningful signals.

Summary:

While Davis offers better conditions than colder locations for PLS-based analysis, it may still face limitations, particularly in months where chill accumulation remains consistently high, reducing variability in the data.

### Sfax, Tunisia

In Sfax, temperature conditions are generally more favorable for PLS regression, particularly between December and February:

-   **December–February:** Temperatures vary enough to create meaningful chill signals, making these months well-suited for PLS-based analysis.

-   **November and March:** Some chill accumulation occurs, but the proportion of effective chill days is much lower compared to peak winter months.

-   **October:** Temperatures remain consistently too high for chill accumulation, meaning no significant chill-related signal is expected.

Summary:

Among the four locations, Sfax presents the best conditions for PLS regression, especially in mid-winter. However, October shows no variation, and November/March may produce only weak signals, limiting their usefulness in identifying temperature response phases.

## Heat model sensitivity vs. observed temperature

Compared to the Dynamic Model, the Growing Degree Hours (GDH) model exhibits a much simpler temperature response. Any day with a minimum temperature above 4°C contributes to heat accumulation, with warmer temperatures leading to a stronger response.

-   Beijing (December–May): Heat accumulation days are rare in December, January, and February, but become frequent from March onward.

-   Other locations: Heat accumulation conditions are met in all months, making it easier to detect response phases.

Across the four sites, there are few months where PLS regression would struggle to identify heat accumulation phases. This likely explains why PLS regression has been more effective at detecting ecodormancy (heat accumulation) phases than endodormancy (chill accumulation) periods, where temperature conditions are often more stable and less variable.

## `Exercises` on expected PLS responsiveness

1.  *Produce chill and heat model sensitivity plots for the location you focused on in previous exercises.*

```{r}
# Generate model sensitivity data
Model_sensitivities_Yakima <-
  Chill_model_sensitivity(latitude = 46.6,
                          temp_models = list(Dynamic_Model = Dynamic_Model,
                                             GDH = GDH),
                          month_range = c(10:12, 1:5))

write.csv(Model_sensitivities_Sfax,
          "Yakima/Model_sensitivities_Yakima.csv",
          row.names = FALSE)

```

```{r warning=FALSE}
# Chill sensitivity model for Yakima

Chill_sensitivity_temps(Model_sensitivities_Yakima,
                        Yakima_weather,
                        temp_model = "Dynamic_Model",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        legend_label = "Chill per day \n(Chill Portions)") +
  ggtitle("Chill model sensitivity at Yakima, USA") 
  
```

```{r warning=FALSE}
# Heat sensitivity model for Yakima

Chill_sensitivity_temps(Model_sensitivities_Yakima,
                        Yakima_weather,
                        temp_model = "GDH",
                        month_range = c(12, 1:5),
                        legend_label = "Heat per day \n(GDH)") +
  ggtitle("Heat model sensitivity at Yakima, USA") 
  
```
